---
title: "ggplot2 exercies"
author: "SASAP - LW"
date: "November 29, 2017"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(tidyr)
  library(dplyr)
})
```

get data: downloading directly from site only works on macs
```{r}
permits <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Aa3c58bd6-481e-4c64-aa93-795df10a4664",
                    stringsAsFactors = FALSE)
```

```{r}
permits = read.table("clipboard", header=T, sep="\t")
```

```{r}
names(permits)
summary(permits)
```

ggplot, dataset, aesthetics funciton (), call a function for the geometry of the plot
start with aesthetics, then supply goemetry instead of making data available to geometry, otherwise 


```{r}
ggplot(permits, aes(Value, StdDev)) + geom_point()

#not as good as:
ggplot(permits) + geom_point(data = permits, aes(Value, StdDev))
```

```{r}
ggplot(permits, aes(Value, StdDev, color = Gear)) + 
  geom_point()
```
aesthetics are not quoted, they are being mapped.
setting is being done

```{r}
ggplot(permits, aes(Value, StdDev, color = "blue")) +
  geom_point()

ggplot(permits, aes(Value, StdDev)) +
  geom_point( color = "blue")

```


```{r}
ggplot(permits, aes(Value, StdDev, color = Gear, size = Range)) + 
  geom_point()

```

```{r}
ggplot(permits, aes(Value)) + 
  geom_histogram()

ggplot(permits, aes(Value)) + 
  geom_histogram(binwidth = 10000)
```

```{r}
ggplot(permits, aes(Gear, Value)) + 
  geom_boxplot()
```


```{r}
#find out how data is named
summary(permits$Region)
unique(permits$Region)

permits_CI = permits %>%
  filter(Region ==  "Cook Inlet")

ggplot(permits_CI, aes(Gear, Value)) +
  geom_violin() + 
  geom_point(pch = 1, position = "jitter")

#but overplotting, so make empty circles, jitter
```

aes (x, y) in plot

```{r}
ggplot(permits, aes(Value, StdDev)) +
  geom_point() +
  geom_smooth(method = lm)
```



```{r}
se_seine = permits %>%
  filter(Gear == "Purse Seine", 
         Region == "Southeast")
summary(se_seine)
```

plus signs pass on data, max is not a ggplot function

```{r}
ggplot(se_seine, aes(Year, Value)) + 
  geom_point() +
  geom_line() + 
  scale_y_continuous(limits = c(0, NA)) + 
  scale_x_continuous(breaks = c(seq(1985, 2015, by = 5)))

```


```{r}
ggplot(se_seine, aes(Year, Value)) + 
  geom_point() +
  geom_line() + 
  coord_cartesian(xlim = c(2000, max(se_seine$Year)))
```

```{r}

valueformat = function(val){
  format(val, scientific = FALSE)
  }

ggplot(permits, aes(Value, StdDev, color = Gear)) + 
  geom_point() + 
  scale_x_continuous(labels = valueformat) + 
  scale_color_manual(values = c("red", "orange","yellow","green", "blue", "purple", "violet"))

ggplot(permits, aes(Value, StdDev, color = Gear)) + 
  geom_point() + 
  scale_x_continuous(labels = valueformat) + 
  scale_color_brewer(palette = "Dark2")



ggplot(permits, aes(Value, StdDev, color = Gear)) + 
  geom_point() + 
  scale_x_log10(labels = valueformat) + 
  scale_color_brewer(palette = "Dark2") + 


```



plotting small multiples using facet
```{r}
ggplot(permits_CI, aes(Year, Value)) + 
  geom_point() + 
  geom_line() + 
  facet_wrap(~Gear,ncol=1)

ggplot(permits_CI, aes(Year, Value)) + 
  geom_point() + 
  geom_line() + 
  facet_wrap(~Gear,dir = "v")
```


```{r}

theme_set(theme_classic())

#customize plot using theme ()
my_theme = theme_classic(base_size = 12) +
  theme(legend.position = "top") 

nice_plot = ggplot(permits, aes(Value, StdDev, color = Gear))+
  geom_point() +
  my_theme
  
```

Save plots using ggsave (others not likely to work)
Default saves the last plot generated
```{r}
ggsave("my_plot.png")
```

```{r}
ggsave("my_plot.png", width = 6, height = 6)
  
```

saving images to a folder
add a directory, decide where to save 
```{r}
ggsave("figures/my_plot.png", width = 6, height = 6)
  
```


```{r}
print(nice_plot)
```

Check out google search results for ggplot extensions, such as 
- ggrepel (labels points so labels do not overlap)
- ggExtra (marginal density plots)
- gganimate (in rmarkdown)


Bonus material on how to include different types of plots to plot a few on a single package: cowplot
plot_grid(p1, p2, align = "hv" or "h" or "v", ncol = 1)
where p1 and p2 are plots





